---
title: "Homework 3: Basic Data Types"
author: "Your Name"
date: "Due: February 17, 2023"
output: html_document
categories: hw
---

[Download the starter qmd file here](https://raw.githubusercontent.com/srvanderplas/unl-stat151/main/homework/03-data-types.qmd)

## Pythagorean Theorem in R

$$a^2 + b^2 = c^2$$

Use the chunk below to define variables $a$, $b$, and $c$, where you set $a$ and $b$ and $c$ is determined mathematically. You may need to find the square root function in R to make this work.

```{r pythagorean-theorem-r}
# Your pythagorean theorem code goes here
```


## Quadratic Formula in Python

$$x = -b \pm \frac{\sqrt{b^2 - 4ac}}{2a}$$


Use the chunk below to define variables $a$, $b$, and $c$, and then define $x1$ and $x2$, the possible values for $x$.

```{python quad-formula-py}
# Your quadratic formula code goes here

```


## Debugging practice

Look at the data frame below, which contains 3 columns: `Name`, `Salary`, and `Hire.Year`. You've read in this data from the [data file online](https://srvanderplas.github.io/stat151-homework/3-data.tsv) and ideally, you want to plot `Hire.Year` on the x-axis and `Salary` on the y-axis (code to do this is provided).

Unfortunately, the data isn't reading in properly. Identify the problem, propose a solution, and try to implement your solution using functions on the R reference card.

Some things you may want to consider:

- `gsub`
- You can access columns in a data frame using `$`: `df$col1` will get you `col1` of the `df` object. (We will talk about data frames more next week, but for now, just try to use them)
- You can access items in a vector using `[]`: `x[1]` will get you the first item in `x`

Note: You should be able to **identify** the problem given what you know about R from this chapter. You may not be able to implement a solution without some additional research, so if you don't get that far that's ok.

```{r}
# This line reads in a tab-separated data file with a header naming the columns
salary <- read.table("https://srvanderplas.github.io/unl-stat151/homework/3-data.tsv",
                     sep = "\t", header = T)

# This code plots labels for each individual at location (x, y)
library(ggplot2)
ggplot(salary, aes(x = Hire.Year, y = Salary, label = Name)) +
  geom_label()

# This displays the salary object
salary
```

-----------------

The problem with the data is:


Here is how I solved the problem:

```{r cleaning-the-data-r}
# Your implementation code goes here

```


