---
title: "Final Exam"
author: "Your Name"
format: html
---

# Instructions

```{r setup, include = F}
# This chunk sets some options for the file compilation
knitr::opts_chunk$set(collapse = T, message = F, warning = F, error = T)
```

For each of these problems, you may choose to solve the problem in either R or python.

Your submitted qmd file must compile without errors (5 points). Use `error=TRUE` in a chunk if it is supposed to return an error (for instance, if you are demonstrating error handling). 

You may use the textbook, your notes, and google on this exam, but you may not post this exam and ask for help on any site, and you must be able to explain how any code you submit on this exam works. It is ok to google, for instance, how to convert a string to a list of characters, but it is not ok to google how to solve the entire question. Please ask if you are concerned about any possible edge cases.

# Scrabble!

Here are the point values for each letter in Scrabble.

- 0 Points - Blank tile.
- 1 Point - A, E, I, L, N, O, R, S, T and U.
- 2 Points - D and G.
- 3 Points - B, C, M and P.
- 4 Points - F, H, V, W and Y.
- 5 Points - K.
- 8 Points - J and X.
- 10 Points - Q and Z.

### Part 1: Functions

(10 points + 3 possible bonus)

Write a function named `scrabble_score(x)` which will take a word as an argument and determine the point value of that word. Assume that the blank tile is indicated by _.

Make sure your function includes input validation, so that if someone calls `scrabble_score(3)`, your function provides an appropriate error.

For a bonus point, make sure your function accepts lower-case and upper-case input and treats both the same.

For +2 points, make your function vectorized, so that you can handle either single characters or a vector of character input.

```{r functions}

```

```{r function-test, error = T, collapse = T}
scrabble_score(3)
scrabble_score("elephant")
scrabble_score("qi")
```

```{python functions-py}

```

```{python function-test-py, error = T, collapse = T}
scrabble_score(3)
scrabble_score("elephant")
scrabble_score("qi")
```


### Scrabble scoring

(15 points)


Read in the text file [scrabble.txt](https://raw.githubusercontent.com/srvanderplas/stat151-homework/main/scrabble.txt) and create a data frame with the word in one column. Create (and fill) a column for the word score and another column with the word length. (3 points per column)

Plot word length and word score in an appropriate plot (`geom_jitter` may be helpful). Describe the plot's main features in 2-3 sentences. Make sure your plot has an appropriate title and appropriate axis labels. Include any annotations (trend lines, etc.) that you think are helpful when describing your plot. (3 points for the plot, 3 for the description)
