---
title: "Building Blocks"
format: revealjs
theme: ./libs/unl/inverse.scss
includes:
  in_header: ./libs/unl/header.html
lib_dir: ./libs
---

## Variables

- Declare: set aside a space in memory for a variable

- Assign: Set the value of a variable

In modern languages (R, Python, etc.) these two steps are usually done simultaneously

`x <- 3` or `y = "hello world"` in R and Python respectively

## Naming Things

- Variable names can't
  - start with a number or symbol
  - include `.` (Python only)
  - be a reserved word
  
- Choose descriptive names
  - Match the equation (if math)
  - May want to include units `length_mm`
  - Avoid single character names where possible    
  `c` in particular is a bad name in R!
  

## Basic Variable Types

- Integer - 2, 6, 109243
- Double - 3.1415, 2.7, 9.0
- String - "ABCDE", " ", "Eddie is a Jack Russell Terrorist"
- Boolean - True, False


## Testing Types

Use a function to test if the type of `y` is `xxx

R: `is.xxx(y)`

Python: `isInstance(y, xxx)` 

## Implicit Type Conversions

R and Python will both try to work with types that don't match... to a point

```{r}
v1 <- 2L + 2.0
v1
is.integer(v1)
is.double(v1)

v2 <- TRUE + 3L
v2
is.logical(v2)
is.integer(v2)
is.double(v2)
```

Variables are converted to the least-complicated/least memory intensive compatible type.

## Implicit Type Conversions

R and Python will both try to work with types that don't match... to a point

```{python}
v1 = 2 + 2.0
v1
isinstance(v1, int)
isinstance(v1, float)

v2 = True + 3
v2
isinstance(v2, bool)
isinstance(v2, int)
isinstance(v2, float)
```

Variables are converted to the least-complicated/least memory intensive compatible type.

## Explicit Type Conversions

You can force things to be converted from one type to the other as well.

```{python}
int("3")
int(3.14)
int(3.84)

bool(3.0)
```

```{r}
as.integer("3")
as.integer(3.14)
as.integer(3.84)

as.logical(3.0)
```
