---
title: "Introduction to Graphics"
format: revealjs
theme: ./libs/unl/inverse.scss
includes:
  in_header: ./libs/unl/header.html
lib_dir: ./libs
execute:
  echo: true
  error: true
---


## We read...

::: columns

::: column

![[Data Visualiztion Basics](https://srvanderplas.github.io/stat-computing-r-python/part-wrangling/02-basic-data-vis.html)](images/datavis-basics-chapter.png){fig-alt="Screenshot of the data visualization basics chapter of the textbook, found at https://srvanderplas.github.io/stat-computing-r-python/part-wrangling/02-basic-data-vis.html."}

:::


::: column


![[Data Visualiztion](https://srvanderplas.github.io/stat-computing-r-python/part-wrangling/02b-graphics.html)](images/datavis-full.png){fig-alt="Screenshot of the data visualization basics chapter of the textbook, found at https://srvanderplas.github.io/stat-computing-r-python/part-wrangling/02b-graphics.html."}

:::

:::


## Syntax basics {.large}

::: columns

::: column


```
ggplot(data = <DATA>) +
  <GEOM>(mapping = aes(<MAPPINGS>), 
         position = <POSITION>,
         stat = <STAT>) +
  <FACET> +
  <COORD> +
  <THEME>
```
:::

::: column

```
import seaborn as sns
import matplotlib.pyplot as plt

plot = sns.<plottype>(<MAPPINGS>)

plt.show() # Show the plot (for quarto)

```
:::

:::

## Practice

```{r}
#| fig-width: 8
#| fig-height: 3
#| echo: false


library(palmerpenguins)
library(ggplot2)
data(penguins)

ggplot(penguins, aes(x = bill_length_mm, fill = species)) + geom_histogram() + facet_wrap(~species) + guides(fill = "none") + xlab("Bill Length, mm") + ylab("Count")
```
  - What type of plot is it? 
  - Which variables are mapped to x, y, and fill? 
  - What is the main message of the plot: what is your main finding, i.e. what do you want viewers to learn from the plot? 
  - Are there any anomalies or outliers?
  - (Extra question) Could you make this plot more effective? How?
  


## Practice

::: columns

::: column

```{r}
#| fig-width: 4
#| fig-height: 4
#| echo: false


library(palmerpenguins)
library(ggplot2)
data(penguins)

ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, color = species)) + geom_point()  + xlab("Bill Length, mm") + ylab("Bill Depth, mm") + theme(legend.position = "bottom") + scale_color_discrete("Species")
```

:::

::: column
  - What type of plot is it? 
  - Which variables are mapped to x, y, and fill? 
  - What is the main message of the plot: what is your main finding, i.e. what do you want viewers to learn from the plot? 
  - Are there any anomalies or outliers?
  - (Extra question) Could you make this plot more effective? How?
  
:::

:::

## Homework

-   Use github classroom to accept the assignment
-   Don't forget committing and pushing the changes!
