---
title: "Homework 9: FizzBuzz Returns!"
author: "Your Name"
format: html
---

## Part 1

Write a FizzBuzz function named `fizzbuzz1`. Your function should take a single number and **return** (not print out) a single string evaluating the number that was passed in. Use input validation to ensure that you are given a number, and provide a useful error if you are not given a number as the function argument.

(Do this in R or python - just pick a language for this assignment. If you choose python, just change the chunk type below.)

```{r fizzbuzz-1}

# Your code goes here



```

This code will test your fizzbuzz function. Please make sure it runs as you expect it to run before you submit this file to Canvas!

```{r testing-fizzbuzz-1, error = T, collapse = T}
fizzbuzz1(3)
fizzbuzz1(5)
fizzbuzz1(7)
fizzbuzz1(15)
fizzbuzz1(0)
fizzbuzz1("this should fail")
```


## Part 2

Write a FizzBuzz function named `fizzbuzz` that can handle a vector of numbers and will **return** (not print out) a vector of character/string values. Check to ensure the input vector is numeric.

Hint: You might use a loop and call your function from the previous part. 

```{r fizzbuzz-2}


```

This code will test your fizzbuzz function. Please make sure it runs as you expect it to run before you submit this file to Canvas!

```{r testing-fizzbuzz-2, error = T, collapse = T}
fizzbuzz(c(0, 3, 5, 7, 11, 15))

fizzbuzz(month.name) # this should fail

```


